### 2020-02-03 thesis/paper.tex
more

\part{Implemented Tool}
\chapter{Using the Tool}\label{cha:Tool}
%maybe not "this chapter..."
This chapter describes what a user can do with the newly implemented tool.

\section{I/O and Data Visualization}
bla

@@

The starting window, in which clusterings can also be loaded back into from the meta-view, also allows for exporting data to CSV. When doing this, the first column in the resulting file will contain the cluster identifiers starting with one. This way computed results can be exported and used in other tools or saved for later.

@@

Data manipulation can be performed as a first step in the starting window to create data-sets or modify them such that they can be used for clustering or visualization. It is also of relevance when loading clustering results back into the starting view to analyze results using dimensionality reduction techniques. For this step three interfaces are of interest:

@@

\subsubsection*{Adding Data Manipulation Methods}
To additional Methods they must be implemented using the corresponding interface and added to the class StartWindow. To add them to this class the static methods initGenerators(), initReducers() and initNormalizers() must add the new method to the list of methods. As an example adding a new generator (newGeneratorX) would require adding ``generators.add(new newGeneratorX());'' to the function initGenerators().

@@

\subsubsection{Consensus Functions}
The consensus functions implement the logic for combining the base clusterings. For this the interface IConsensusFunction is of interest and can be extended to add new functions. Per default two main implementations are available. Both implementations allow for the user to either define a target value for the number of resulting clusters $k$ within each result, or for the algorithm to guess this parameter itself. Firstly, CoAssociationMatrixAverageLink can be used, which computes a Co-Association (CA) Matrix and combines objects via the average link strategy. When $k$ is defined the algorithm joins together sets of points until only $k$ sets remain, which is used as final answer. If no $k$ is defined the algorithm performs two runs, calculating the lifetime \cite{lifetime} of the resulting tree and choosing the level to cut as the one resulting with maximum lifetime. As second method DICLENS is available, which was generously provided by Mimaroglu and Aksehirl i\cite{DISCLENS}.

@@

To visualize the data in different ways and provide a functional Graphical User Interface (GUI) the Java Swing \cite{javaswing} Toolkit was used. It allows for both high level calls creating objects withing windows, e.g. panels, buttons or combo-boxes, as well as low level calls by directly changing the behavior of the draw() function. For the basic GUI the panels and buttons of the tool were created with high level calls, while the different graphs and plots were implemented by overriding the draw() function. Overriding the draw() function immensely improved performance as the already available high level objects are computationally costly by comparison, when many of them are needed .

\subsection{Scatter Plot}

>>>>

\part{Implemented Tool}
\chapter{Using the Tool}\label{cha:Tool}
%maybe not "this chapter..."
The implemented tool aims to facilitate the computation and selection of good clustering solutions for a given data-set. To do this both powerful visualizations and algorithms are needed, which is combined in this tool. The usual workflow using the tool can be described with three steps, each with its own view.

\section{Data Visualization and Manipulation}
Firstly the user needs to import or create a data-set on which the clustering should be performed. When the application is started a scatter plot without data is shown, as well as a side menu with different buttons.

%%image

 To import a data-set

@@

The starting window or data-view, in which clusterings can also be loaded back into from the meta-view, also allows for exporting data to CSV. When doing this, the first column in the resulting file will contain the cluster identifiers starting with one. This way computed results can be exported and used in other tools or saved for later.

@@

Data manipulation can be performed as a first step in the data-view to create data-sets or modify them such that they can be used for clustering or visualization. It is also of relevance when loading clustering results back into the data-view to analyze results using dimensionality reduction techniques. For this step three interfaces are of interest:

@@

\subsubsection*{Adding Data Manipulation Methods}
To additional Methods they must be implemented using the corresponding interface and added to the class DataView. To add them to this class the static methods initGenerators(), initReducers() and initNormalizers() must add the new method to the list of methods. As an example adding a new generator (newGeneratorX) would require adding ``generators.add(new newGeneratorX());'' to the function initGenerators().

@@

\subsubsection{Consensus Functions}
The consensus functions implement the logic for combining the base clusterings. For this the interface IConsensusFunction is of interest and can be extended to add new functions. Per default two main implementations are available. Both implementations allow for the user to either define a target value for the number of resulting clusters $k$ within each result, or for the algorithm to guess this parameter itself. Firstly, CoAssociationMatrixAverageLink can be used, which computes a Co-Association (CA) Matrix and combines objects via the average link strategy. When $k$ is defined the algorithm joins together sets of points until only $k$ sets remain, which is used as final answer. If no $k$ is defined the algorithm performs two runs, calculating the lifetime \cite{lifetime} of the resulting tree and choosing the level to cut as the one resulting with maximum lifetime. As second method DICLENS is available, which was generously provided by Mimaroglu and Aksehirl \cite{DICLENS}.

@@

To visualize the data in different ways and provide a functional Graphical User Interface (GUI) the Java Swing \cite{javaswing} Toolkit was used. It allows for both high level calls creating objects withing windows, e.g. panels, buttons or combo-boxes, as well as low level calls by directly changing the behavior of the draw() function. For the basic GUI the panels and buttons of the tool were created with high level calls, while the different graphs and plots were implemented by overriding the draw() function. Overriding the draw() function immensely improved performance as the already available high level objects are computationally costly by comparison, when many of them are needed. This section only includes additional information on the implementation and functionality, information on the usage of the visual elements in the tool can be found in Section [REFERENCE].

\subsection{Scatter Plot}
The scatter plot is the simplest visualization of the data within the tool. It is implemented though two axis and a canvas object. The axis include logic for the draw the value range and selected dimension, including the mouse listeners for modifying it, as well as the translation from model coordinates to viewer coordinates. The canvas implements the drawing of data points, including different shapes and colors depending on their cluster identifier, whether or not they are filtered out or if they represent special objects like the ground truth in the meta-view. Additionally the canvas can support mouse listeners enabling interactivity with the points.
%more?
